NanoMaLy

### 2015-11-18 paper/pldi16-submission/related.tex

\section{Related Work}
\label{sec:related-work}

\subsection{Type Error Localization}
\label{sec:type-error-local}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%

\begin{itemize}
\item Lerner \etal~\cite{lerner_searching_2007} attempt to localize the
  type error and suggest a fix by replacing expressions (or removing
  them entirely) with alternatives based on the surrounding program
  context.
\item Chen and Erwig~\cite{chen_counter-factual_2014} use a variational
  type system represent the possibility that the inferred type of an
  expression may be incorrect, because the expression may be the source
  of the error. It then attempts to deduce the error source by searching
  for an expression whose type can be changed such that type inference
  would succeed. In contrast to Seminal, which searches for changes at
  the value-level, by searching at the type level the search is complete
  due the finite universe of types in a program.
\item Neubauer and Thiemann~\cite{neubauer_discriminative_2003} present
  a decidable type system based on discriminative sum types, in which
  all terms are typeable and type derivations contain all type errors in
  a program. They then use the typing derivation to slice out the parts
  of the expression related to each error.
\item Zhang and Myers~\cite{zhang_toward_2014} present an algorithm for
  identifying the most likely culprit in a system of unsatisfiable
  constraints (\eg type equalities), based on Bayesian reasoning.
\item Pavlinovic \etal~\cite{pavlinovic_finding_2014} translate the
  error localization problem to a MaxSMT optimization problem, using
  compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}
\end{itemize}

@@

@@

\subsection{Misc}
\begin{itemize}
\item \cite{vytiniotis_equality_2012} extend the Haskell compiler GHC to
  support compiling ill-typed programs, but their intent is rather
  different from ours. Their goal was to allow programmers to
  incrementally test refactorings, which often cause type errors in
  distant pieces of code. They replace any expression that fails to
  type check with a \emph{runtime} error, but do not perform
  any type checking at runtime.
\item \cite{perera_functional_2012}
\end{itemize}

>>>>

\section{Related Work}
\label{sec:related-work}
In this section we connect our work to related efforts on type errors,
testing, and program exploration.

\subsection{Type Errors}
\label{sec:type-error}

\paragraph{Diagnosis and Repair}
\label{sec:diagnosis-repair}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%
Thus, many groups have explored techniques to pinpoint the true source
of the error and recommend fixes.

Lerner \etal~\cite{lerner_searching_2007} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{chen_counter-factual_2014} use a variational type
system represent the possibility that the inferred type of an expression
may be incorrect, because the expression may be the source of the
error. It then attempts to deduce the error source by searching for an
expression whose type can be changed such that type inference would
succeed. In contrast to Seminal, which searches for changes at the
value-level, by searching at the type level the search is complete due
the finite universe of types in a program.
%
Neubauer and Thiemann~\cite{neubauer_discriminative_2003} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{zhang_toward_2014} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{pavlinovic_finding_2014} translate the error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis \etal~\cite{vytiniotis_equality_2012} extend the Haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant pieces of code. They replace any expression that fails to type
check with a \emph{runtime} error, but do not perform any type checking
at runtime.

@@

\subsection{Program Exploration}
Perera \etal~\cite{perera_functional_2012} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} -- a value containing holes -- and in turn present
a partial slice, containing only those steps that affected the specified
components of the partial value. This system is designed to answer
questions of the form ``Where did this value come from?'' and thus is
focused on backward exploration. In contrast, our trace visualization
supports forward \emph{and} backward exploration, though our backward
steps are more limited than Perera's. Specifically, we do not support
selecting a value and inserting the intermediate terms that preceded it
while ignoring unrelated computation steps; this would be interesting
future work.

\ES{todo: more on program exploration}
